---
title: "Stat 3301: Homework 1"
author: "Nathan Johnson.9254"
date: "Due by date and time specified on Carmen"
output: pdf_document
---

\

Setup:
```{r message=FALSE}
library(alr4)
library(tidyverse)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. Use the option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.

\

#### Question 1
Import the data set `ames_real_estate.csv` (which is available on Carmen) and use it to do the
following:

a) Find the smallest and largest observed sale prices separately for each 
neighborhood.
b) Make a scatterplot with the square footage of the house on the *x*-axis 
and the sale price of the house on the *y*-axis.
c) Write a sentence describing the relationship between `SalePrice` and `SqFt`.

\

##### Solution to Question 1
```{r}
library(readr)
ames = read_csv('ames_real_estate.csv')
```
a. Here are the smallest and largest observed sale prices separately for 
each neighborhood:
```{r}
ames %>% group_by(Neighborhood)%>% summarize(minSalePrice = min(SalePrice), 
                          maxSalePrice = max(SalePrice)) %>% print(n = Inf)
```

b. Here is the scatterplot of ames with square feet on the x-axis and sale price
on the y-axis.
```{r}
ames %>% ggplot(aes(x = SqFt, y = SalePrice)) + geom_point() + theme_bw(16)
```
c. As square feet (`SqFt`) in a house increases on average, the sales price 
(`SalePrice`) of the house also increases on average.

\

#### Question 2
Use the data set `Heights` from the library`alr4` to calculate the average 
height for daughters whose mothers are **at least** 60 inches tall. 
(The purpose of this question is to make sure you have the package installed 
correctly.)

\

##### Solution to Question 2
```{r}
Heights %>% filter(mheight >= 60) %>% summarize(mean(dheight))
```
The mean height of daughters whose mothers were at least 60 inches tall is 
64.09704 inches.
