---
title: "Stat 3301: Homework 5"
date: "Nathan Johnson"
author: "Due by date and time specified on Carmen"
output: pdf_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will

* use plots of data to assess whether log transforms are useful.
* see an example of the "log rule" in practice.
* fit regression models with the predictor and/or response in log scale.
* interpret $\hat{\beta}_1$ when the predictor and/or response are in log scale.
* construct confidence intervals when the predictor is in log scale.
* make a plot of a fitted model when the predictor is in log scale.

------

#### Question 1
<!--
...
-->
The data set `ufcwc` in the `alr4` package contains data on a sample of western cedar trees
that was taken in 1991 in the Upper Flat Creek stand of the University of Idaho
Experimental Forest. Each case in the data set represents a tree. In this question we will 
look at the relationship between the variables:

| Variable      | Description                                                    |
| ------------- | -------------------------------------------------------------- |
| `Height`      | height of the tree in decimeters                               |
| `Dbh`         | diameter of the tree in mm measured at 137 cm above the ground |

We will be interested in using tree diameter (`Dbh`) to predict tree height (`Height`).

(a) For the linear regression model $Height_i = \beta_0 + \beta_1 \log Dbh_i + e_i$, report
the numeric values of estimated parameters $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\sigma}^2$.

(b) Use the fitted model to compute "by hand" (i.e., don't use the `predict` function) 
a 95\% confidence interval for average tree height for trees that have a diameter of 500 mm.

(c) Make a scatterplot with `log(Dbh)` on the x-axis and `Height` on the y-axis, and add
the estimated linear regression line to the plot.

(d) Make a scatterplot with `Dbh` (on its original scale) on the x-axis and `Height` on 
the y-axis. Use your fitted model from part (a) to add a (curved) line to the plot that represents 
the estimated average tree height as a function of tree diameter, $\hat{E}(\mathtt{Height} \mid 
\mathtt{Dbh})$. Add dashed (curved) lines to the plot that represent point-wise, 95\% confidence
intervals for the mean.

(e) Using the estimated slope coefficient, $\hat{\beta}_1$, write a sentence that quantifies
what our model says about how average tree height depends on tree diameter (on its original scale).


\


##### Solution to Question 1
Your answers go here.

```{r}
library(alr4)
```
A.
```{r}
X = ufcwc$Dbh
Y = ufcwc$Height
SlogXY = sum((log10(X) - mean(log10(X)))*(Y))
SlogXX = sum((log10(X) - mean(log10(X)))^2)

B1 = SlogXY / SlogXX
B0 = mean(Y) - (B1 * mean(log10(X)))
lm.ufc = lm(Height ~ log10(Dbh), data = ufcwc)

s2 = sum((Y - (B0 + B1*log10(X)))^2)/(length(X)-2)
B1
B0
s2
```
$\hat{\beta}_0 = -463.3, \hat{\beta}_1 = 275.2, \hat{\sigma}^2 = 1111.19$

b.
Use $\hat{\beta}_0 \pm t(\alpha/2, n-2) * se(\hat{\beta}_0\mid X)$

```{r}
value = B0 + B1 * log10(500)
alpha = 0.05
n = length(ufcwc$Height)
mean = mean(ufcwc$Dbh)
se = sqrt(s2) * sqrt((1/n) + (log10(500) - mean(log10(ufcwc$Dbh)))/SlogXX)
lwr = value - pt(alpha/2, n - 2, lower.tail = FALSE) * se
uwr = value + pt(alpha/2, n - 2, lower.tail = FALSE) * se
lwr
uwr
```
The average tree height will be between 276.58 and 282.32 decimeters with 95% confidence.

c.
```{r}
plot(log10(ufcwc$Dbh), ufcwc$Height, xlab = "log Diameter of Tree in mm", ylab = "Height of Tree", main = "Trees")
abline(a = B0, b = B1, col = "Red")
```
d.
```{r}
d.grid = data.frame(Dbh=seq(from=min(ufcwc$Dbh), to=max(ufcwc$Dbh), length.out=100))
ypred = predict(lm.ufc, newdata=d.grid, interval="confidence")
plot(ufcwc$Dbh, ufcwc$Height, pch=16, xlab="Dbh in mm", ylab="Height in dm",
main="Mean Fuction", ylim=c(0,500))
lines(d.grid[,1], ypred[,1], col="blue", lwd=2)
lines(d.grid[,1], ypred[,2], col="blue", lty=2); lines(d.grid[,1], ypred[,3], col="blue", lty=2)
```

e.
```{r}
TenPercent = B1 * log10(1.1)
TenPercent
```

For every 10% larger the diameter of the tree is, there will be a 11.39 decimeter increase in height of the tree.





#### Question 2
<!--
...
-->
The data set `AdRevenue.txt` on `Carmen` (Source: Sheather, 2009) contains data
on the `n = 70` top magazines in the US in terms of gross advertising revenue in
2006. The variables in the data set include the magazine name, the parent company 
of the magazine, the total ad revenue per page (in thousands of dollars) and
the magazine circulation (in millions). This problem concerns predicting gross
advertising revenue per page in 2006 from circulation.


(a) Make four plots: (i) `AdRevenue` vs. `Circulation`, (ii) `AdRevenue` vs. `log(Circulation)`, (iii) `log(AdRevenue)` vs. `Circulation`, (iv) `log(AdRevenue)` vs. `log(Circulation)`.
Explain why transforming both `AdRevenue` and `Circulation` to log scale is useful for building
a linear regression model that uses circulation to predict ad revenue.

(b) Recall the book's definition of the "log rule" (p188): "If the values of a variable range 
over more than one order of magnitude and the variable is strictly positive, then replacing 
the variable by its logarithm is likely to be helpful" in linearizing relationships. Is that the 
case here? Say why or why not.

(c) A company is interested in quantifying how increases in circulation are associated with changes
in ad revenue. To do this, they fit a linear regression model with `log(Circulation)` as the
predictor and `log(AdRevenue)` as the response. Use the estimated parameter $\widehat{\beta}_1$ 
to provide, in plain English, a careful interpretation of how a change in the predictor 
variable `Circulation` (on its original scale) is associated with a change in average ad 
revenue (on its original scale).


\


##### Solution to Question 2
Your answers go here.

a. 
```{r}
rev = read.csv('AdRevenue-1.txt')
plot(rev$Circulation, rev$AdRevenue, xlab = "Circulation of magazine", ylab = "Ad Revenue in Millions of Dollars")
plot(log10(rev$Circulation), rev$AdRevenue, xlab = "log Circulation of magazine", ylab = "Ad Revenue in Millions of Dollars")
plot(rev$Circulation, log10(rev$AdRevenue), xlab = "Circulation of magazine", ylab = "log Ad Revenue in Millions of Dollars")
plot(log10(rev$Circulation), log10(rev$AdRevenue), xlab = "log Circulation of magazine", ylab = "log Ad Revenue in Millions of Dollars")
```
Plots i, ii, iii are all not suitable for linear modeling. It is easy to see just from visuals that there is not a linear relation between these models. But plot iv does appear to be linear.

b.
Yes, it is the case here. While said rule applies to plot ii and plot iii, it unfortunately did not seem to create the linear model we were hoping for. But plot iv does create what appears to be a linear model.

c.
```{r}
SlogXlogYrev = sum((log10(rev$Circulation) - mean(log10(rev$Circulation)))*(log10(rev$AdRevenue))) / (length(rev$Circulation) - 1)
SlogXXrev = sum((log10(rev$Circulation) - mean(log10(rev$Circulation)))^2) / (length(rev$Circulation) - 1)

B1rev = SlogXlogYrev / SlogXXrev
B1rev

interpret = exp((log10(1.1))*B1rev) - 1
interpret
```

For a 10% increase in circulation, there is a 2.21% increase in ad revenue.









